---
title: "Assignment 5"
output: html_document
---

# Setup

This document requires the following packages installed, `caret`, `rpart`, `tree`, `randomForest`, `e1071` and `ggplot`.

```{r loading_libraries}
library(ggplot2)
library(rpart)
library(caret)
library(randomForest)
library(e1071)
```

# Loading and summarizing the data

We start the assignment by loading the SeaFlow data from the `seaflow_21min.csv` file:

```{r seaflowLoad, cache=TRUE}
seaflow <- read.csv("seaflow_21min.csv")
```

Let's look at a summary of the data:

```{r summary, cache=TRUE}
summary(seaflow)
```

# Creating the test set

Now we create a training set:

```{r partition, echo=FALSE, cache=TRUE}
samples <- createDataPartition(seaflow$cell_id, 2)
trainingSet <- seaflow[samples$Resample1,]
testSet <- seaflow[samples$Resample2,]
```

Now let's calculate the mean for `time`:

```{r mean_of_time, cache=TRUE}
mean(trainingSet$time)
```

# Plotting `pe` against `chl_small`

```{r pe_to_chl_small_plot, cache=TRUE}
qplot(pe, chl_small, data = seaflow, color=pop)
```

# Training the decision tree

```{r training_tree, cache=TRUE}
fol <- formula(pop ~ fsc_small + fsc_perp + fsc_big + pe + chl_big + chl_small)
model <- rpart(fol, method="class", data=trainingSet)
print(model)
```

# Predicting the values with decision trees

```{r predicting_with_decision_trees, cache=TRUE}
dtPredictions <- predict(model, newdata=testSet, type="class")
sum(testSet$pop == dtPredictions)
sum(testSet$pop == dtPredictions)/nrow(testSet)
```

# Predicting the values with random forests

```{r predicting_with_random_forests, cache=TRUE}
randomForestModel <- randomForest(fol, data=trainingSet)
rfPredictions <- predict(randomForestModel, newdata=testSet, type="class")
sum(testSet$pop == rfPredictions)
sum(testSet$pop == rfPredictions)/nrow(testSet)
importance(randomForestModel)
```

# Predicting with support vector machines

```{r predicting_with_svm, cache=TRUE}
svmModel <- svm(fol, data=trainingSet)
svmPredictions <- predict(svmModel, newdata=testSet, type="class")
sum(testSet$pop == svmPredictions)
sum(testSet$pop == svmPredictions)/nrow(testSet)
```

# Building confusion matrices

```{r confusion_matrices, cache=TRUE}
table(pred = dtPredictions, true = testSet$pop)
table(pred = rfPredictions, true = testSet$pop)
table(pred = svmPredictions, true = testSet$pop)
```

# plotting data to find out which one is broken

```{r plots, cache=TRUE}
qplot(chl_big, file_id, data = seaflow, color=pop)
qplot(chl_small, file_id, data = seaflow, color=pop)
qplot(pe, file_id, data = seaflow, color=pop)
qplot(fsc_big, file_id, data = seaflow, color=pop)
qplot(fsc_perp, file_id, data = seaflow, color=pop)
qplot(fsc_small, file_id, data = seaflow, color=pop)
```

# Finding out a time series data error

```{r, time_and_chl_big, cache=TRUE}
qplot(time, chl_big, data = seaflow, color=pop)
```

# Removing data from the broken file

```{r subsetting, cache=TRUE}
cleaned <- subset(seaflow, file_id != 208)
cleanedSamples <- createDataPartition(cleaned$file_id, 2)
cleanedTrainingSet <- cleaned[cleanedSamples$Resample1,]
cleanedTestSet <- cleaned[cleanedSamples$Resample2,]
cleanedSvmModel <- svm(fol, data=cleanedTrainingSet)
cleanedSvmPredictions <- predict(cleanedSvmModel, newdata=cleanedTestSet, type="class")
```

# Showing the prediction results

```{r clean_prediction_results, cache=TRUE}
c(nrow(cleanedTestSet),length(cleanedSvmPredictions))
c(nrow(testSet),length(svmPredictions))
sum(cleanedTestSet$pop == cleanedSvmPredictions)
sum(cleanedTestSet$pop == cleanedSvmPredictions)/nrow(cleanedTestSet)
sum(testSet$pop == svmPredictions)/nrow(testSet)
sum(cleanedTestSet$pop == cleanedSvmPredictions)/nrow(cleanedTestSet) - sum(testSet$pop == svmPredictions)/nrow(testSet)
```